Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-09 csf-2017/paper.tex
f3 text

For both programs, we can prove that if they are linked up with an
adversary that is allowed to allocate new memory but otherwise has no
capabilities, then executing the program, the assertion will never
fail.
We can only prove this, because the capability machine encapsulates
the local state of the two programs. The two examples also illustrate
the versatility of the logical relation. The two examples use
different calling convention, but the logical relation is not specific
to any calling convention, so we can use it to reason about both
programs. In order to reason about the unknown untrusted part, we use
the Fundamental Theorem to argue that it must be capability safe.
\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f1: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[]$)
    pop $r_1$
    assert $r_1$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a stack-based calling convention.}
  \label{fig:prog-f1}
\end{figure}

\todo[inline]{give a taste of a proof by giving high-level proof here?
Maybe later for one of the other examples}

@@

\subsection{Well-bracketed control-flow}

The program $f3$ in \figurename~\ref{fig:prog-f3}
\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

>>>>

For both programs, we can prove that if they are linked up with an
adversary that is allowed to allocate new memory but otherwise has no
capabilities, then executing the program, the assertion will never
fail.
We can only prove this, because the capability machine encapsulates
the local state of the two programs. The two examples also illustrate
the versatility of the logical relation. The two examples use
different calling convention, but the logical relation is not specific
to any calling convention, so we can use it to reason about both
programs. In order to reason about the unknown untrusted part, we use
the Fundamental Theorem to argue that it must be capability safe.

\todo[inline]{give a more detailed high-level taste of a proof here? Maybe later for one of the other examples?}

\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f1: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[]$)
    pop $r_1$
    assert $r_1$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a stack-based calling convention.}
  \label{fig:prog-f1}
\end{figure}

@@

\subsection{Well-bracketed control-flow} 
Using the stack-based calling convention of \texttt{scall}, we get
well-bracketed control-flow. To illustrate this, we look at
two example programs \texttt{f3} in \figurename~\ref{fig:prog-f3} and
\texttt{g1} in \figurename~\ref{fig:prog-g1}.

In \texttt{f3} there are two calls to an adversary and in order for
the assertion in the middle to succeed, they need to be
well-bracketed. If the adversary is able to store the protected return
pointer, so it is available for the second call, then it can be used
in a context where \texttt{f3} has 2 on top of its stack which will cause the
assertion to fail. The protected return pointer is passed as a local
capability, so the only means the adversary has to store it is the
stack capability. The stack capability itself is a local capability,
so the adversary can only store it on the stack. This means that after
the adversary jumps to the protected return pointer, \texttt{f3} can
revoke it (by overwritting it which happens as part of the
\texttt{scall}). In other words, the adversary cannot store the
protected return pointer for the second call, and it is possible to
prove that the assertion never fails.
\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

Our final example is \texttt{g1} which is a faithful translation of
what is known as the ``awkward'' example.
\todo[inline]{insert references}
